General constraints on magic distillation protocols from statistical mechanics

### 2021-06-02 magic_paper.tex
Wrote basic proof of bounds, left comment on how to extend. Checked D_\infty

2. If $D_\infty (\rho,\tau) = 0$ this implies that $\max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)}=1$ and the initial slope of $L_{\rho|\tau}(x)$ is $1$. From the concavity of the function this implies that the slope of $L_{\rho|\tau}(x)$ must equal $1$ throughout the interval $[0,1]$ and this, together with the definition of the Lorenz curve, implies that $\frac{W_\rho(\x)}{W_\tau(\x)}=1$ for all $\x$. Since the Wigner representation $\rho \mapsto W_\rho(\x)$ is bijective this implies that $\rho = \tau$ only. The converse holds by inspection.

3. Given a vector $\w \in \mathbb{R}^N$ in general it is the case that
\begin{equation}
\max_k w_k \ne \max_k |w_k|.
\end{equation}
However we do have that
\begin{equation}
\max_{k_1,k_2} (w_{k_1} w_{k_2}) = \max_k |w_k|^2 = \max_k w_k^2.
\end{equation}
If $\w \in \mathbb{R}^N$ with $w_k \ge 0$ for all $k$ then
\begin{equation}
\max_{k_1, k_2, \dots, k_n} [ w_{k_1}w_{k_2}\cdots w_{k_n}] = (\max_k w_k)^n,
\end{equation}
and therefore for arbitrary $\w \in \mathbb{R}^N$ and $n$, we have
\begin{align}
\hspace{-1cm}\max_{k_1, k_2, \dots, k_{2n}} [ w_{k_1}w_{k_2}\cdots w_{k_{2n}}] &\le \max_{k_1, k_2, \dots, k_{2n}}[|w_{k_1}w_{k_2}|\cdots |w_{k_{2n-1}} w_{k_{2n}}|] \nonumber \\
&=(\max_{k_1 ,k_2} |w_{k_1} w_{k_2}| )^n \nonumber \\
&\le\max_{k_1, k_2, \dots, k_{2n}} [ w_{k_1}w_{k_2}\cdots w_{k_{2n}}]  .
\end{align}
and so we have that in general
\begin{equation}
\max_{k_1, k_2, \dots, k_{2n}} [ w_{k_1}w_{k_2}\cdots w_{k_{2n}}]  =  (\max_{k_1 ,k_2} [w_{k_1} w_{k_2}] )^n.
\end{equation}
If we now let $\w = W_{\rho|\tau}(\x) := W_{\rho}(\x) /W_{\tau}(\x)$ we have
\begin{equation}
\max_{\x_1, \x_2, \dots ,\x_{2n}} \prod_{k=1}^{2n} W_{\rho|\tau}(\x_k) = (\max_{\x_1, \x_2} W_{\rho|\tau}(\x_1)W_{\rho|\tau}(\x_2))^n
\end{equation}
and therefore taking logarithms we have that $D_\infty (\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ as required. \ddd{[Check details for any fine print.]}

@@

Using the above results, we can show that $H_{2m}$ is Schur-concave for all quasi-distributions.
In particular, for any non-negative integer $m$, we have that
\begin{equation}
	W_{\rho_n}(\bmx) \succ W_{\rho'_{n'}}(\bmx) \mbox{ implies } H_{2m}(\rho_n) \leq H_{2m}(\rho'_{n'}).
\end{equation}
By additivity of $H_{2m}$, we get $n H_{2m}(\rho) \leq n' H_{2m}(\rho')$, or equivalently,
\begin{equation}
	\frac{n'}{n} \geq \frac{H_{2m}(\rho)}{H_{2m}(\rho')}.
\end{equation}

\nick{Wrong way? I don't think that $m \rightarrow \infty$ leads to our bound}

\ddd{[TO-DO AND TIDY AND CHECK]} Okay great...this looks easy to do, we just need to compute the terms
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_k} (w_1, w_2, \dots, w_N),
\end{equation}
for Renyi entropies (which are symmetric), and look to see when the above inequality holds. 
We assume $ \alpha=2m$ is an even integer, so that $H_\alpha (\w)$ is well-defined on quasi-distributions. We then have that
\begin{align}
V_k:=-\frac{\partial H_\alpha}{\partial w_k} &= \partial_k \frac{1}{\alpha-1} \log \sum_i w_i^{2m} \\
&= \frac{1}{(2m-1) \sum_i w_i^{2m}}2mw_k^{2m-1}.
\end{align}
We now impose the monotonicity condition $V_1 \ge V_2 \ge V_3 \cdots \ge V_N$. This requires, setting $S= \sum_i w_i^{2m}$ for simplicity, that
\begin{align}
 \frac{1}{(2m-1)S}2mw_i^{2m-1}\ge \frac{1}{(2m-1)S}2mw_{i+1}^{2m-1},
\end{align}
for each $i=1,2,\dots, N-1$. We assume that $m >0$ and $2m \ne 1$ and so this becomes
\begin{equation}
w_i^{2m-1}\ge w_{i+1}^{2m-1}\mbox{ for all } 1 \le i < N.
\end{equation}
For $2m >1$ this condition holds whenever $w_i \ge w_{i+1}$, which is true in $\D^\downarrow$. Therefore $-H_{2m}(\w)$ is Schur-Convex on quasi-distributions for any integer $m\in \mathbb{N}$. This implies it is a monotone function for majorization, namely,
\begin{equation}
\w \succ \w' \mbox{ implies } H_{2m} (\w) \le H_{2m} (\w') \mbox{for any } m \in \mathbb{N}.
\end{equation}
This therefore should give \emph{additional} distillation bounds in the unital fragment. We'd need to check additivity is okay, but this should be fine. Therefore
\begin{equation}
W_{\rho_n}(\x) \succ W_{\rho'_m}(\x) \mbox{ implies } n H_{2m}(\rho) \le m H_{2m}(\rho'). 
\end{equation}
(Some sign error here? Should get:)
\begin{equation}
R \le \frac{H_{2m}(\rho)}{H_{2m}(\rho')}.
\end{equation}
\ddd{[Plan: check all these calculations! Give the generalised unital bound. Then we just conjecture this extends to Renyi-divergences and leave it at this.]}
\subsection{Dropping pre/post-processing of the magic states}
\label{app:cliff_processing}

>>>>

2. $D_\infty (\rho,\tau) = 0$ implies that $\max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)}=1$ and the initial slope of $L_{\rho|\tau}(x)$ is $1$. From the concavity of the function this implies that the slope of $L_{\rho|\tau}(x)$ must equal $1$ throughout the interval $[0,1]$ and this, together with the definition of the Lorenz curve, implies that $W_\rho(\x)/W_\tau(\x) = 1$ for all $\x$. Since the Wigner representation $\rho \mapsto W_\rho(\x)$ is bijective this implies that $\rho = \tau$ only. The converse holds by inspection.

3. Given a vector $\w \in \mathbb{R}^N$, it is generally the case that $\max_k w_k \ne \max_k |w_k|$. 
However, we do have that
\begin{equation}
	\max_{k_1,k_2} (w_{k_1} w_{k_2}) = \max_{k_1,k_2} |w_{k_1} w_{k_2}|,
\end{equation}
and if additionally $w_k \geq 0$ for all $k$, then we also have
\begin{equation}
	\max_{k_1, k_2, \dots, k_n} \left( w_{k_1}w_{k_2}\cdots w_{k_n} \right) = (\max_k w_k)^n.
\end{equation}
Therefore, for arbitrary $\w \in \mathbb{R}^N$ and $n$, we have
\begin{align}
	&\max_{k_1, k_2, \dots, k_{2n}}\left( w_{k_1}w_{k_2}\cdots w_{k_{2n}} \right) = \nonumber\\
	&\max_{k_1, k_2, \dots, k_{2n}}\left( |w_{k_1}w_{k_2}|\cdots |w_{k_{2n-1}}w_{k_{2n}}| \right) = \nonumber \\
	&\left( \max_{k_1 ,k_2}\left( w_{k_1} w_{k_2} \right) \right)^n.
\end{align}

If we now let $\w = W_{\rho|\tau}(\x) := W_{\rho}(\x) /W_{\tau}(\x)$ we have
\begin{equation}
\max_{\x_1, \x_2, \dots ,\x_{2n}} \prod_{k=1}^{2n} W_{\rho|\tau}(\x_k) = \left( \max_{\x_1, \x_2} W_{\rho|\tau}(\x_1)W_{\rho|\tau}(\x_2) \right)^n
\end{equation}
and therefore taking logarithms we have that $D_\infty (\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ as required.

@@

\ddd{[TO-DO AND TIDY AND CHECK]} Okay great...this looks easy to do, we just need to compute the terms
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_k} (w_1, w_2, \dots, w_N),
\end{equation}
for Renyi entropies (which are symmetric), and look to see when the above inequality holds. 
We assume $ \alpha=2m$ is an even integer, so that $H_\alpha (\w)$ is well-defined on quasi-distributions. We then have that
\begin{align}
V_k:=-\frac{\partial H_\alpha}{\partial w_k} &= \partial_k \frac{1}{\alpha-1} \log \sum_i w_i^{2m} \\
&= \frac{1}{(2m-1) \sum_i w_i^{2m}}2mw_k^{2m-1}.
\end{align}
We now impose the monotonicity condition $V_1 \ge V_2 \ge V_3 \cdots \ge V_N$. This requires, setting $S= \sum_i w_i^{2m}$ for simplicity, that
\begin{align}
 \frac{1}{(2m-1)S}2mw_i^{2m-1}\ge \frac{1}{(2m-1)S}2mw_{i+1}^{2m-1},
\end{align}
for each $i=1,2,\dots, N-1$. We assume that $m >0$ and $2m \ne 1$ and so this becomes
\begin{equation}
w_i^{2m-1}\ge w_{i+1}^{2m-1}\mbox{ for all } 1 \le i < N.
\end{equation}
For $2m >1$ this condition holds whenever $w_i \ge w_{i+1}$, which is true in $\D^\downarrow$. Therefore $-H_{2m}(\w)$ is Schur-Convex on quasi-distributions for any integer $m\in \mathbb{N}$. This implies it is a monotone function for majorization, namely,
\begin{equation}
\w \succ \w' \mbox{ implies } H_{2m} (\w) \le H_{2m} (\w') \mbox{for any } m \in \mathbb{N}.
\end{equation}

\ddd{[Plan: check all these calculations! Give the generalised unital bound. Then we just conjecture this extends to Renyi-divergences and leave it at this.]}

Using the above results, we can derive the following entropic bound on magic state transformations. \nick{to be extended (1) from unital processes to general relative majorization, (2) from $\alpha = 2m$ to $\alpha = 2m/(2k+1)$ for integer $m,k \geq 0$, (1) from Strange state to any magic state distillation}

\begin{theorem}
	Consider the magic distillation process
	\begin{equation}
		\rho_{\rm{S}}(\epsilon)^{\otimes n} \rightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'},
	\end{equation}
	in the unital fragment.
	Define the $\alpha$-Renyi entropy of a state $\rho$ as
	\begin{equation}
		H_{\alpha}(\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha.
	\end{equation}
	Then, we have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho_{\rm{S}}(\epsilon) \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho_{\rm{S}}(\epsilon') \right)},
	\end{equation}
	for $\alpha = 2m,\ m\in\mathbb{N}$.
\end{theorem}
\begin{proof}
	We can regularise the number of output copies in the process by tensoring in $(n-n')$ copies of the (free) maximally mixed state $\id/3$, so that we instead consider the process
\begin{equation}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \rightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')}.
\end{equation}
The entropy $H_\alpha$ is well-defined on the input and output states for $\alpha = 2m$ and since we have that $W_{\rho_n}(\bmx) \succ_{\id/3} W_{\rho'_{n'}}(\bmx)$, by Schur-concativity of $H_\alpha$ this implies
\begin{equation}
	H_\alpha \left( \rho_{\rm{S}}(\epsilon)^{\otimes n} \right) \leq H_\alpha \left( \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')} \right).
\end{equation}
By additivity of $H_\alpha$, we can rewrite the inequality as
\begin{equation}
	n H_\alpha \left( \rho_{\rm{S}}(\epsilon) \right) \leq n' H_\alpha \left( \rho_{\rm{S}}(\epsilon') \right) + (n - n') H_\alpha \left( \id/3 \right).
\end{equation}
Now, $H_\alpha \left( \id/3 \right)$ is the maximum value of $H_\alpha(\rho)$ for unital processes because every state $\rho$ majorizes $\id/3$, so we can finally re-arrange the inequality to
\begin{equation}
	\frac{n'}{n} \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho_{\rm{S}}(\epsilon) \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho_{\rm{S}}(\epsilon') \right)}.
\end{equation}
\end{proof}

\subsection{Dropping pre/post-processing of the magic states}
\label{app:cliff_processing}
